---
title: "AIPTW Simulation Study — PLR vs Cox Outcome Model"
subtitle: "Apr 28 protocol, Section 6"
author: "AIPTW workstream"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    df-print: paged
    embed-resources: true
    fig-width: 9
    fig-height: 5
execute:
  echo: false
  warning: false
  message: false
params:
  results_dir: "../results"
---

```{r setup}
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
})

# Paths -- params$results_dir is relative to the .qmd location.
results_dir <- params$results_dir
summary_path <- file.path(results_dir, "summary.csv")
raw_dir      <- file.path(results_dir, "raw")
truth_dir    <- file.path(results_dir, "truth")

if (!file.exists(summary_path)) {
  stop("Summary CSV not found at ", summary_path,
       ". Run summarise_aiptw.R first.")
}

summary_df <- read.csv(summary_path, stringsAsFactors = FALSE)
raw_files  <- list.files(raw_dir, pattern = "\\.csv$", full.names = TRUE,
                          recursive = FALSE)
raw_df     <- do.call(rbind, lapply(raw_files, read.csv,
                                     stringsAsFactors = FALSE))
truth_files <- list.files(truth_dir, pattern = "\\.csv$", full.names = TRUE,
                           recursive = FALSE)
truth_df    <- do.call(rbind, lapply(truth_files, read.csv,
                                      stringsAsFactors = FALSE))

# Defaults for older outputs without the method column.
if (is.null(raw_df$method))     raw_df$method     <- "aiptw_plr"
if (is.null(summary_df$method)) summary_df$method <- "aiptw_plr"
```

# Overview

This report consolidates the AIPTW simulation results from
`results/raw/<scenario>.csv` and the protocol metrics in
`results/summary.csv`. It covers up to two AIPTW variants depending on
how `simulate_aiptw.R` was invoked:

- **`aiptw_plr`** — discrete-time pooled-logistic outcome model with
  canonical logit link; the protocol-faithful AIPTW formulation per
  the methods document Section 4.2.2.
- **`aiptw_cox`** — Cox PH outcome model wrapping
  `riskRegression::ate()`. Same propensity-score model, different
  outcome specification.

```{r overview-table}
overview <- summary_df %>%
  group_by(method, dgm) %>%
  summarise(
    n_scenarios = n_distinct(scenario_id),
    n_replicates = mean(n_reps_total),
    n_failed     = sum(n_reps_failed),
    .groups = "drop"
  )

kable(overview, caption = "Run scope by method and DGM",
      digits = 0)
```

# Section 6.5 — Performance metrics

The Apr 28 protocol's primary deliverable. One row per
`(method, scenario, t, target)`. RD power is reported only for the RD
target; coverage and SE error are reported for all three targets.

```{r perf-table}
perf <- summary_df %>%
  filter(t == max(summary_df$t)) %>%
  select(method, dgm, misspec, rho_L, target,
         truth, mean_est, rel_bias_pct,
         empirical_sd, mean_model_se, rel_se_err,
         coverage, power, n_reps_ok)

kable(perf, digits = 4,
      caption = sprintf("Performance at t = %g (largest evaluation time)",
                        max(summary_df$t)))
```

# Section 7 — Bias and coverage by scenario

```{r bias-fig, fig.cap="Relative bias (%) at each evaluation time, by misspecification scenario."}
bias_plot <- summary_df %>%
  filter(target %in% c("S0", "S1", "RD")) %>%
  ggplot(aes(x = factor(t), y = rel_bias_pct,
             fill = method)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.85) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  facet_grid(target ~ dgm + misspec, scales = "free_y") +
  theme_classic(base_size = 9) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(size = 7),
        strip.text  = element_text(size = 7)) +
  labs(x = "t (years)", y = "Relative bias (%)",
       fill = "Method")

bias_plot
```

```{r coverage-fig, fig.cap="Coverage (proportion of 95% bootstrap CIs containing truth)."}
cov_plot <- summary_df %>%
  ggplot(aes(x = factor(t), y = coverage,
             fill = method)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.85) +
  geom_hline(yintercept = 0.95, linetype = "dashed", colour = "red") +
  facet_grid(target ~ dgm + misspec, scales = "free_y") +
  ylim(0, 1) +
  theme_classic(base_size = 9) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(size = 7),
        strip.text  = element_text(size = 7)) +
  labs(x = "t (years)", y = "Coverage",
       fill = "Method")

cov_plot
```

# Survival curves vs analytic Weibull truth

For each scenario, the figure overlays the replicate-mean estimated
S^a(t) curves (one per method) on the analytic Weibull truth. With the
protocol-grade R = 1900 these mean curves should overlap the truth
under the doubly-robust scenarios.

```{r curves-prep}
# Replicate-mean S(t) curves at each evaluation time (protocol grid only).
curve_df <- raw_df %>%
  filter(target %in% c("S0", "S1"), status == "ok") %>%
  group_by(method, scenario_id, dgm, misspec, rho_L, t, target) %>%
  summarise(mean_est = mean(est, na.rm = TRUE), .groups = "drop") %>%
  rename(A = target) %>%
  mutate(A = ifelse(A == "S0", 0L, 1L))

truth_long <- bind_rows(
  truth_df %>% transmute(scenario_id, dgm, misspec, rho_L, t,
                          A = 0L, surv = S0),
  truth_df %>% transmute(scenario_id, dgm, misspec, rho_L, t,
                          A = 1L, surv = S1)
) %>%
  mutate(method = "True Weibull")

curves <- bind_rows(
  curve_df %>% transmute(method, scenario_id, dgm, misspec, rho_L,
                          t, A, surv = mean_est),
  truth_long
) %>%
  mutate(A = factor(A, levels = c(0L, 1L),
                    labels = c("Control (A=0)", "Treatment (A=1)")))
```

```{r curves-fig, fig.cap="Replicate-mean S^a(t) by method (PLR / Cox / Truth), faceted by scenario.", fig.height=8}
curves %>%
  ggplot(aes(x = t, y = surv, colour = method, linetype = method)) +
  geom_line(linewidth = 0.6) + geom_point(size = 1) +
  facet_grid(misspec ~ dgm + A) +
  ylim(0, 1) +
  theme_classic(base_size = 9) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 7)) +
  labs(x = "t (years)", y = "Marginal Survival Probability")
```

# PLR vs Cox AIPTW — head-to-head

Side-by-side comparison of the two AIPTW outcome-model choices on the
same scenarios.

```{r plr-vs-cox}
if (length(unique(summary_df$method)) >= 2L) {
  pivot_cols <- c("rel_bias_pct", "coverage", "mean_ci_width")
  comp <- summary_df %>%
    select(method, dgm, misspec, rho_L, t, target,
           rel_bias_pct, coverage, mean_ci_width) %>%
    pivot_wider(names_from = method,
                values_from = all_of(pivot_cols))
  kable(comp, digits = 4,
        caption = "Per-cell comparison: PLR vs Cox AIPTW")
} else {
  cat("Only one AIPTW variant present in this run (",
      paste(unique(summary_df$method), collapse = ", "),
      "). Re-run with --method both to populate this table.")
}
```

# Bootstrap vs influence-function inference (Cox arm only)

```{r boot-vs-if}
cox_rows <- summary_df %>%
  filter(method == "aiptw_cox", n_if_ok > 0)

if (nrow(cox_rows) > 0L) {
  cox_compare <- cox_rows %>%
    select(dgm, misspec, rho_L, t, target,
           empirical_sd,
           mean_model_se, rel_se_err, coverage,
           mean_if_se, rel_if_se_err, if_coverage)
  kable(cox_compare, digits = 4,
        caption = "Cox-AIPTW: bootstrap-based vs IF-based SE / coverage")
} else {
  cat("No Cox-AIPTW rows with IF-based SEs available.")
}
```

# Failure audit

```{r failure-audit}
fail_df <- summary_df %>%
  filter(n_reps_failed > 0 |
         (!is.na(mean_n_boot_failed) & mean_n_boot_failed > 0))

if (nrow(fail_df) > 0L) {
  kable(
    fail_df %>%
      select(method, scenario_id, t, target,
             n_reps_total, n_reps_ok, n_reps_failed,
             mean_n_boot_failed, boot_errors_seen),
    digits = 1,
    caption = "Cells with non-zero replicate or bootstrap failure"
  )
} else {
  cat("No replicate or bootstrap failures detected.")
}
```

# Run reproduction

```{r repro}
cat("To reproduce this report:\n\n")
cat("```bash\n")
cat("# AIPTW-PLR arm\n")
cat("Rscript run_aiptw_plr_analysis.R --R 1900 --B 200 --workers 16\n\n")
cat("# AIPTW-Cox arm\n")
cat("Rscript run_aiptw_cox_analysis.R --R 1900 --B 200 --workers 16\n\n")
cat("# Aggregate\n")
cat("Rscript summarise_aiptw.R\n\n")
cat("# This report\n")
cat("quarto render reports/AIPTW_report.qmd\n")
cat("```\n")
```

# Session info

```{r session}
sessionInfo()
```
